\chapter{Future Work}

While the results of our proposed methodology and tooling are promising, there are limitations. Constraints were necessary to manage the scope of work for a PhD thesis, and others for technical reasons. While the CyclicL tool is self-contained with regards to features and requirements, there is no traceability between the \gls{UCD}, the goal diagram, and the tool for \tool. This presents challenges to comprehensive traceability from requirements through the features to their sources. This also causes a gap in analysis when performing iterative changes as there are no facilities currently that can enable change impact analysis between feature modelling and the \gls{UCD}. As users learn more about their product we anticipate and encourage them to iterate through the various phases of development in this methodology and we want to support that computationally as well.

Using this methodology to support safety and security assurance case development is another avenue we would like to explore. As this methodology emphasizes traceability, there could be opportunities for supporting assurance case development. This possibility is part of the reason for the design decisions to have the \texttt{Safety} type of requirement in the \tool\ metamodel. This would include the traceability to the eventual modelling environments for \gls{UCD}s and goal diagrams in a tool for \tool.

Another interesting avenue of research for the PACEMAKER requirements was completed by Vivien Wang et al.~\cite{wang2012identification} explores the assumptions made in the PACEMAKER requirements document~\cite{scientific2007pacemaker}. A potential avenue of future work would be to integrate the process of analysis for identifying assumptions made in system requirements with the \tool\ methodology to allow for traceability to assumptions. This would be a nice to have in order to ensure that assumptions made during development are identified, documented, and traceable throughout the product lifecycle.

%This presents another challenge of traceability and justification. It requires the engineer to be aware of the links between the domain and problem space analysis and the feature and requirement models in the tool. This implicit knowledge will eventually need to be documented somewhere to support both future development changes and assurance for safety-critical development. As a result, one potential path is creating a bridge between the domain analysis modelling and the current tool to extend the traceability all the way up to the concept phase of development. While we expect \gls{UCD} and goal diagrams to remain mostly static once they are created, the direct traceability could be helpful documentation for onboarding new employees to help them get familiar with why certain decisions were made, support change impact analysis justifications, and aid the development of assurance case by having complete end-to-end traceability between the problem space and the solution space.

The implementation of the tool for \tool\ is still incomplete as we have only demonstrated the capability of the minimum viable product. There are many functions we will need to add to have a more complete tool. There are currently no constraints in the tool to ensure that all user created models are valid according to the metamodel. While the tool offers flecibility, there is space for enhanced validation to provide a more robust analysis support to users who are creating models. It is, after all, a trade-off between flexibility and error checking.

We are currently relying on conventions and good engineering to create semantically meaningful models but this is very difficult to maintain at scale within a company or industry. Another lane of development would be in allowing users to create these models textually as well as graphically. Very often adoptability of a tool, especially an \gls{MDE} tool is difficult as many engineers are unfamiliar with \gls{MDE} techniques and approaches. Creating a textual environment for creating these models may help with usability and approachability of CyclicL for use in industry. Next, we would want to implement a method of consistency checking within the tool. This would remove a lot of the burden from the engineer to ensure that their features are consistent and their requirement are consistent between features. This is a unique challenge as the requirements are specified using the Gherkin behavioural approach and thus may require us to re-evaluate what type of specification language we implement within the requirement canvases of CyclicL.

Other future work tasks that have been mentioned throughout this thesis include integration with existing test compilers, integration with previous feature modelling tools, and general UI enhancements. As there are many Gherkin test environments, building a bridge to allow integration with \tool\ would be helpful to verification tracking and traceability. Integration with another feature modelling tool, such as FeatureIDE, would be serve as an exploration for the impact of the feature-requirement hierarchy defined in this thesis on existing feature weaving and composition techniques.


